---
title: "STAT/MATH 495: Problem Set 01"
author: "Sarah Teichman"
date: "2017-09-07"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    collapsed: false
    smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=8, fig.height=4.5)

# Load packages
library(tidyverse)
library(mosaic)
```

Downloading data
```{r include=FALSE}
#download data
train <- read_csv("https://rudeboybert.github.io/STAT495/static/Titanic/train.csv")
test <- read_csv("https://rudeboybert.github.io/STAT495/static/Titanic/test.csv")
#submission <- read_csv("https://rudeboybert.github.io/STAT495/static/Titanic/gender_submission.csv")
```

Here I looked at variables that might be useful, including Sex, Age, and Cabin.
```{r}
tally(train$Sex)
summary(train$Age)
train <- mutate(train, cabinB = ifelse(is.na(Cabin),0,1))
tally(train$cabinB)
#making cabinB, an indicator variable for whether or not the person had a cabin (seems to be highly correlated with the Pclass variable, but with only two levels)
tally(Survived~Sex,data=train)
tally(Survived~cabinB,data=train)
ggplot(train,aes(Age,fill=factor(Survived)))+geom_histogram(binwidth=10,position="dodge")
```

Here I created datasets with only women and only men to explore relationships between variables. 

```{r}
women <- filter(train,Sex=="female")
men <- filter(train,Sex=="male")
tally(Survived~cabinB,data=women)
tally(Survived~cabinB,data=men)
women0 <- filter(women,cabinB==0)
men1 <- filter(men,cabinB==1)
ggplot(women0,aes(x=Age,fill=factor(Survived)))+geom_histogram(binwidth=8,position="fill")
ggplot(men1,aes(x=Age,fill=factor(Survived)))+geom_histogram(binwidth=8,position="fill")

```

From the plot I can see that the percentage of women without cabins who survived was greater than 50% for almost every age group, so in my submission I predict that all of the women survived. However, cabinB is a better indicator of men who survived in the training set, and within that group the youngest and oldest men survived. In my submission I predict that all of the men died except for the ones who had cabins and were either younger than 20 or older than 75.

```{r}
test <- mutate(test,survived=0,cabinB= ifelse(is.na(Cabin),0,1))
x <- which(test$Sex=="female")
y <- which(test$Sex=="male"&test$cabinB==1&(test$Age<20|test$Age>75))
test$survived[y] <- 1
test$survived[x] <- 1
ID <- test$PassengerId
submission <- data.frame(PassengerID=ID,Survived=test$survived)
tally(~Survived,data=submission,format="prop")
tally(~Survived,data=train,format="prop") 
#in my predictions, there is a similar percentage of people that survived as in the training set
write.csv(file="submission.csv", x=submission,row.names=F)
```